Optimizing DTW-Based Audio-to-MIDI Alignment and Matching

### 2016-08-18 paper.tex
Fix minor notational issues found while writing thesis

To create this dataset, we first collected 1,000 MIDI files which were transcriptions of Western popular music songs.
We then applied the following series of transformations, based on our experience with common differences between MIDI transcriptions and audio recordings:
First, to simulate differences in tempo, we adjusted the timing in each MIDI file by a low-frequency length-$N$ random signal $r$, defined as
\begin{align*}
s[k] &\sim \mathcal{N}(0, \sigma_t), k \in \{0, \ldots, N\}\\
R[k] &= \begin{cases}
s[k]e^{-k}, &k \in \{0, \ldots, N\}\\
R[2N - k], &k \in \{N + 1, \ldots, 2N - 1\}
\end{cases}\\
r[n] &= \sum_{k = 0}^{2N - 1} R[k]e^{j\pi kn/N}, n \in \{0, N - 1\}
\end{align*}
i.e.\ the inverse discrete Fourier transform of an exponentially decaying Gaussian-distributed random spectrum with standard deviation $\sigma_r$.
Second, the first 10\% and last 10\% of the transcription were each cropped out with probability 50\%, which simulates the MIDI file being an incomplete transcription.
In addition, 1\% of each transcription was cut out at a random location with 10\% probability, which simulates a missing measure.
Third, because it is common for a MIDI transcription to be missing an instrument (for example, karaoke versions of songs, in which the lead vocal line has been deleted, are frequently distributed as MIDI files), we removed each instrument track in each MIDI file with probability $p_r$, making sure never to remove all instruments.
Fourth, in all MIDI files, we randomly added 1 or -1 to the program number of each instrument.
This simulated the fact that when comparing a synthesized MIDI file to an audio recording, the timbre of a synthesized MIDI instrument is always somewhat different from its real-world counterpart.
Finally, for each note in each instrument, we multiplied the velocity by a random number sampled from $\mathcal{N}(1, \sigma_v)$ while keeping it in the MIDI velocity range $[1, 127]$.
This was meant to further simulate differences in instrument characteristics in real-world vs.\ synthesized songs, and also simulated missing notes for large $\sigma_v$.
All MIDI data manipulations were realized with \texttt{pretty\char`_midi} \cite{raffel2014pretty_midi}.

@@

Based on the design choices outlined in Section \ref{sec:dtw}, we chose to optimize over the following parameter space:
\begin{description}[topsep=1pt,itemsep=-1pt,leftmargin=5pt]
\item[Feature representation:] We used either chroma vectors or constant-Q spectra.
The constant-Q spectra spanned 4 octaves, starting from MIDI note C3 (65.4 Hz) with 12 bins per octave.
In preliminary experiments, we found that all of the best-performing alignment systems used log-magnitude features regardless of whether chroma vectors or constant-Q spectra were used, so we computed log-magnitude features in all experiments.
\item[Time scale:] We either computed feature vectors every 46 milliseconds or utilized beat-synchronous features.
\item[Normalization:] We optionally z-scored the feature vectors, and normalized them by their $L^1$, $L^2$, $L^\infty$ (max) norm, or not at all.
\item[Penalty:] For the penalty, we optimized a scale in $[0, 3]$ to apply to the median distance between all pairs of feature vectors in $X$ and $Y$.
Using the median distance made this penalty adaptive to different feature representations and normalization schemes.
\item[Gully and band path constraint:] We allowed the gully to take any value in $[0, 1]$ and optionally enforced the band path constraint.
\end{description}
This space subsumes most of the systems discussed in Section \ref{sec:dtw}.
All feature extraction was realized with \texttt{librosa} \cite{mcfee2015librosa, mcfee2015librosa_scipy}.

>>>>

To create this dataset, we first collected 1,000 MIDI files which were transcriptions of Western popular music songs.
We then applied the following series of transformations, based on our experience with common differences between MIDI transcriptions and audio recordings:
First, to simulate differences in tempo, we adjusted the timing in each MIDI file by a low-frequency length-$N$ random signal $r$, defined as
\begin{align*}
s[k] &\sim \mathcal{N}(0, \sigma_t), k \in \{0, \ldots, N\}\\
R[k] &= \begin{cases}
s[k]e^{-k}, &k \in \{0, \ldots, N\}\\
R[2N - k], &k \in \{N + 1, \ldots, 2N - 1\}
\end{cases}\\
r[n] &= \sum_{k = 0}^{2N - 1} R[k]e^{j\pi kn/N}, n \in \{0, \ldots, N - 1\}
\end{align*}
i.e.\ the inverse discrete Fourier transform of an exponentially decaying Gaussian-distributed random spectrum with standard deviation $\sigma_t$.
Second, the first 10\% and last 10\% of the transcription were each cropped out with probability 50\%, which simulates the MIDI file being an incomplete transcription.
In addition, 1\% of each transcription was cut out at a random location with 10\% probability, which simulates a missing measure.
Third, because it is common for a MIDI transcription to be missing an instrument (for example, karaoke versions of songs, in which the lead vocal line has been deleted, are frequently distributed as MIDI files), we removed each instrument track in each MIDI file with probability $p_r$, making sure never to remove all instruments.
Fourth, in all MIDI files, we randomly added 1 or -1 to the program number of each instrument.
This simulated the fact that when comparing a synthesized MIDI file to an audio recording, the timbre of a synthesized MIDI instrument is always somewhat different from its real-world counterpart.
Finally, for each note in each instrument, we multiplied the velocity by a random number sampled from $\mathcal{N}(1, \sigma_v)$ while keeping it in the MIDI velocity range $[1, 127]$.
This was meant to further simulate differences in instrument characteristics in real-world vs.\ synthesized songs, and also simulated missing notes for large $\sigma_v$.
All MIDI data manipulations were realized with \texttt{pretty\char`_midi} \cite{raffel2014pretty_midi}.

@@

Based on the design choices outlined in Section \ref{sec:dtw}, we chose to optimize over the following parameter space:
\begin{description}[topsep=1pt,itemsep=-1pt,leftmargin=5pt]
\item[Feature representation:] We used either chroma vectors or constant-Q spectra.
The constant-Q spectra spanned 4 octaves, starting from MIDI note C2 (65.4 Hz) with 12 bins per octave.
In preliminary experiments, we found that all of the best-performing alignment systems used log-magnitude features regardless of whether chroma vectors or constant-Q spectra were used, so we computed log-magnitude features in all experiments.
\item[Time scale:] We either computed feature vectors every 46 milliseconds or utilized beat-synchronous features.
\item[Normalization:] We optionally z-scored the feature vectors, and normalized them by their $L^1$, $L^2$, $L^\infty$ (max) norm, or not at all.
\item[Penalty:] For the penalty, we optimized a scale in $[0, 3]$ to apply to the median distance between all pairs of feature vectors in $X$ and $Y$.
Using the median distance made this penalty adaptive to different feature representations and normalization schemes.
\item[Gully and band path constraint:] We allowed the gully to take any value in $[0, 1]$ and optionally enforced the band path constraint.
\end{description}
This space subsumes most of the systems discussed in Section \ref{sec:dtw}.
All feature extraction was realized with \texttt{librosa} \cite{mcfee2015librosa, mcfee2015librosa_scipy}.
